### 2022-06-20 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@6e59aed62281889e728b148dd8371076deca2901 🚀

Spillover risk is complex and, even within a pool of susceptible
wildlife hosts, driven by a multiplicity of factors (Plowright et al.
2017). Although host richness is often used as a coarse proxy for
hotspots of emergence risk (Anthony et al. 2017, Ruiz-Aravena et al.
2022), these approaches deliberately oversimplify interspecific
heterogeneity in immunology, behavior, and other traits. Global maps of
spillover risk often struggle to distill these features into
interpretable risk maps, and overlook highly-unique host pools that
allow for the rapid evolution of highly divergent viruses (Agosta et al.
2010). In the case of generalist pathogens like betacoronaviruses, there
is conceptual and empirical support to the idea that these
community-level mechanisms are even more important in driving hotspots
of overall risk (Power and Mitchell 2004).

These kinds of dynamics lend themselves to interpretation through the
lens of the Geographic Mosaic Theory of Coevolution (GMTC; Thompson
2005), which connects microevolutionary dynamics in symbiotic
interactions to macroecological dynamics in host communities. The GMTC
posits that coevolutionary processes between pairs (Thompson 1994) or
complexes (Janzen 1980) of species are structured in space by the
rippling effects of abiotic conditions onto evolutionary mechanism,
resulting in spatially fragmented evolutionary dynamics, coupled only by
dispersal-related processes (Gomulkiewicz et al. 2000). In turn, these
spatially fragmented processes can lead taxonomically homogeneous
systems to have different structure and dynamics over large spatial
extents (see \emph{e.g.} Price 2002). The GMTC predicts a spatial
fragmentation of coevolutionary dynamics under the joint action of three
processes (see notably Gomulkiewicz et al. 2007), which all have the
potential to act on outbreak potential, pathogen tramission, and disease
virulence (Parratt et al. 2016, Turner et al. 2021). First hot- and
coldspots of coevolution can appear when the intensity of
\emph{interaction} (in terms of reciprocal fitness consequences) varies
spatially, because of \emph{e.g.} partial range overlap between
organisms (Nuismer et al. 2003). Second, the GMTC supposes the existence
of selection mosaics, wherein the intensity of \emph{selection} varies
across space; the strength of reciprocal selection responds to the
biotic complexity of the community (locally diverse hosts and viruses
are more biotically complex; Thrall et al. 2007) \emph{and} to the local
favorability of the environment (Hochberg and Baalen 1998). Finally,
trait remixing occurs when coevolutionary dynamics are driven by the
arrival (or departure) of functional traits, through changes in
community composition due to invasions, meta-community dynamics, and
disperal.

Each of these elements can be applied to a relatively quantifiable
aspect of host-virus ecology: (i) viral sharing among hosts,
representing the strength of potential interaction between viruses and
any one host (i.e., places where viruses undergo constant host switching
may be coevolutionary coldspots); (ii) the phylogenetic diversity of
hosts, as a proxy for variation in the immunological mechanisms that
antagonize viruses (i.e., the selection mosaic); and (iii) the local
uniqueness of the bat community, representing the potential for viruses
to be exposed to novel host traits (e.g., variation in receptor
sequences). Together, we argue that these can be used to identify the
evolutionary drivers that---in conjunction with transmission processes
(e.g., viral prevalence in reservoirs and animal-human contact rates)---
determine disease emergence risk.

Here,

We turn the processes on the GMTC into definitions of spillover risk
from viruses to hosts (focusing on the bats-betacoronavirus complex),
with a specific emphasis on the potential to create independent
coevolutionary dynamics (and therefore spatial fragmentation in the
risk) through heterogeneity.

@@

\leavevmode\hypertarget{ref-Hochberg1998AntCoe}{}%
Hochberg, M. E. and Baalen, M. 1998. Antagonistic coevolution over
productivity gradients. - The American Naturalist 152: 620--634.

@@

\leavevmode\hypertarget{ref-Nuismer2003CoeHos}{}%
Nuismer, S. L. et al. 2003. Coevolution between hosts and parasites with
partially overlapping geographic ranges. - Journal of Evolutionary
Biology 16: 1337--1345.

@@

\leavevmode\hypertarget{ref-Parratt2016InfDis}{}%
Parratt, S. R. et al. 2016. Infectious Disease Dynamics in Heterogeneous
Landscapes. - Annual Review of Ecology, Evolution, and Systematics 47:
283--306.

@@

\leavevmode\hypertarget{ref-Thrall2007CoeSym}{}%
Thrall, P. H. et al. 2007. Coevolution of symbiotic mutualists and
parasites in a community context. - Trends in Ecology \& Evolution 22:
120--126.

\leavevmode\hypertarget{ref-Turner2021RolEnv}{}%
Turner, W. C. et al. 2021. The roles of environmental variation and
parasite survival in virulence--transmission relationships. - Royal
Society open science 8: 210088.

>>>>

Disease emergence is complex, and is driven not only by animal-human
contact, but also by the underlying evolutionary dynamics in viral
reservoirs (Plowright et al. 2017). Although host richness is often used
as a superficial proxy for spillover risk (Anthony et al. 2017,
Ruiz-Aravena et al. 2022), these approaches oversimplify the relevant
interspecific heterogeneity in immunology, behavior, and other traits,
and therefore overlook unique host pools that allow for the rapid
evolution of highly divergent viruses (Agosta et al. 2010). In the case
of generalist pathogens like betacoronaviruses, there is conceptual and
empirical support to the idea that these community-level mechanisms are
even more important (Power and Mitchell 2004). This creates a disconnect
between coevolutionary theory and ecological studies of spillover risk.

The Geographic Mosaic Theory of Coevolution (GMTC; Thompson 2005)
attempts to explicitly connect microevolutionary dynamics to the
macroecology and biogeography of symbiotic interactions. The GMTC posits
that coevolutionary processes between pairs (Thompson 1994) or complexes
(Janzen 1980) of species are structured in space by the rippling effects
of abiotic conditions onto evolutionary mechanism, giving rise to
fragmented systems with different structure and ecologically dynamics
over large spatial extents (see \emph{e.g.} Price 2002). The GMTC
predicts a spatial fragmentation of coevolutionary dynamics under the
joint action of three processes (see notably Gomulkiewicz et al. 2007):
coevolutionary hot- and coldspots, which appear when the intensity of
\emph{interaction} (in terms of reciprocal fitness consequences) varies
spatially; selection mosaics, wherein the intensity of \emph{selection}
varies across space, driven by both the biotic complexity of the
community (locally diverse hosts and viruses are more biotically
complex) and the local favorability of the environment
(\textbf{Thrall2007?}); and trait remixing, which occurs when
coevolutionary dynamics are driven by by the arrival (or departure) of
functional traits, through changes in community composition due to
invasions, meta-community dynamics, and disperal.

Here, we develop a framework that quantifies each of these elements in
light of a measurable aspect of host-virus interactions. To do so, we
explore the global biogeography of betacoronaviruses, the group that
includes SARS-CoV, MERS-CoV, and SARS-CoV-2. Working from a recently
published database of bat hosts of betacoronaviruses, we develop the
first global maps of both host and virus evolutionary distinctiveness
and biogeographic regions for this system. Aiming to explain these
patterns, we develop a generalized framework for applying the GMTC to
host-virus interactions, with a specific emphasis on the potential to
create independent coevolutionary dynamics (and therefore spatial
fragmentation in risk) through heterogeneity. We develop a trivariate
risk assessment system that connects each GMTC mechanism o a
quantifiable aspect of host-virus interactions: (i) viral sharing rates
in host communities, representing the strength of potential interaction
between viruses and any one host (i.e., places where viruses undergo
constant host switching may be coevolutionary coldspots); (ii) the
phylogenetic diversity of hosts, as a proxy for variation in the
immunological mechanisms that antagonize viruses (i.e., the selection
mosaic); and (iii) the local uniqueness of the bat community,
representing the potential for viruses to be exposed to novel host
traits (e.g., variation in receptor sequences). Together, we argue that
these can be used to identify and map the evolutionary drivers that---in
conjunction with transmission processes (e.g., viral prevalence in
reservoirs and animal-human contact rates)--- determine disease
emergence risk.

@@

@@

@@

@@
